Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering

### 2016-10-28 nips2016/nips2016.tex
nips v3: submitted version, mostly written by Xavier (2016-05-20)

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Convolutional Neural Networks for Graphs}

@@

\begin{abstract}
This paper introduces Graph Neural Network (GNN), a novel framework to process
data on irregular, i.e.~non-Euclidean domains modelized as graphs. Even recent
developments in Deep Learning who use graph embeddings don't make use of the
graph structure per se.
\end{abstract}

\section{Introduction}\label{introduction}

Large networks allow to learn complex relationships. Although they
overfit easily if the training set is small because there is a lot of
parameters to learn.

Convolutional neural networks (CNN) offer a great reduction of
parameters on regular Euclidean domain. This work proposes to generalize
it to the discrete model of manifolds, graphs. Such an approach was
proposed by \citep{bruna_spectral_2013, henaff_deep_2015}, we will
refine it with proper graph signal processing tools.

This paper introduces a framework to use CNN with signals defined on
weighted graphs. Our work leverages spectral methods, which are much
faster than a spatial approach would be, in the same sense as doing
convolutions as multiplication in the Fourier domain (with the help of
the FFT) is much faster than multiplying patches on an image.

Another option to do clustering or regression on data graphs is to first
embed the graph and then use the embedding as additional features.

\section{Related work}\label{related-work}

Comparison with fully connected NN (Reuters used in
\citep{srivastava_dropout_2014})

\begin{itemize}
\item
  + less parameters via convolution
\item
  ? accuracy
\end{itemize}

Comparison with first generation graph CNN \citep{henaff_deep_2015}.

\begin{itemize}
\item
  + better accuracy
\item
  = same number of parameters
\item
  - graph construction vs supervised estimation
\item
  Theoretical:

  \begin{itemize}
  \item
    + proper spectral graph signal processing tools
  \item
    + kNN graph (sparse weight matrix) vs fully connected graph. SVD
    complexity for Fourier basis is O(n\^{}3), we avoid it.
  \item
    + METIS vs naive agglomeration
  \item
    + Chebyshev polynomials vs splines to approximate filters Filtering
    from \(O(n^2)\) to \(O(|E| K)\) where \(K\) is the polynomial order
    approximately \(O(n)\) if the weight matrix is sparse, i.e.~grows
    linearly with \(N\) This addresses the first limitation mentioned in
    \citep{henaff_deep_2015}.
  \end{itemize}
\end{itemize}

\section{Model}\label{model}

\subsection{Spectral Graph Theory}\label{spectral-graph-theory}

A graph \(\G = (\V, \E, \W)\) is defined by a set \(V\) of \(|V| = M\)
nodes and a set \(E\) of weighted edges. The connectivity of the graph
is captured by the adjacency matrix \(W \in \R^{M \times M}\) which
entry \(W_{i,j}\) denotes the weight of the edge \((v_i, v_j) \in E\)
which connects the vertex \(v_i \in V\) to \(v_j \in V\). It is set to
\(0\) if the vertices are not connected, i.e. \((v_i, v_j) \notin E\).
Assuming an undirected graph, \(W\) is a symmetric matrix. A graph
signal is any signal \(x \in \R^M\) defined on the vertices of \(G\).

The combinatorial graph Laplacian is defined as
\[ \L^c := D - W \in \R^{M \times M}, \] where \(D \in \R^{M \times M}\)
is the diagonal degree matrix defined as \(D_{i,i} := \sum_j W_{i,j}\).
Note that it is a difference operator such that
\[ (\L^c x)_i = \sum_j W_{i,j} (x_i - x_j). \] The normalized graph
Laplacian is defined as
\[ \L^n := I_M - D^{-1/2} W D^{-1/2} \in \R^{M \times M}, \] where
\(I_M \in \R^{M \times M}\) is the identity matrix. Finally, the
random-walk graph Laplacian is defined as
\[ \L^{rw} := I_M - D^{-1} W \in \R^{M \times M}. \] Note that this work
is independent of the chosen graph Laplacian \(\L\), which can be any of
\(\L^c\), \(\L^n\) or \(\L^{rw}\).

In analogy to the real line Fourier Transform, a Fourier basis is given
by the eigenvectors of the Laplacian
\[ \L u_i = \lambda_i u_i \st \norm{u_i}_2=1,
\ u_i \in \R^M, \ i = 0, \ldots, M-1, \] with their associated
eigenvalues \(\lambda_i\). Assuming the graph is connected, we may order
the vector \(\lambda := [\lambda_0, \ldots, \lambda_{M-1}]^T \in \R^M\)
of eigenvalues such that
\[ 0 = \lambda_{min} = \lambda_0 < \lambda_1 \leq \ldots \leq
\lambda_{M-1} = \lambda_{max}. \] As the Laplacian is a real symmetric
and \gls{PSD} matrix, the eigenvalues are real and positive, and the
eigenvectors are orthonormal. The Laplacian is indeed diagonalized by
the Fourier basis \(U := [u_0, \ldots, u_{M-1}] \in \R^{M \times M}\)
such that
\begin{equation} \L = U \Lambda U^T, \label{eq:lap_diag}\end{equation}
where \(\Lambda := \diag(\lambda) \in \R^{M \times M}\) denotes a
diagonal matrix of eigenvalues and \(U^T\) is the matrix transpose of
\(U\). See \citep{chung_spectral_1997} for details on spectral graph
theory.

The Graph Fourier Transform \(\hat{x} \in \R^M\) of any graph signal
\(x\) is given by \[ \hat{x} = U^T x =
[\langle u_0, x \rangle, \ldots, \langle u_{M-1}, x \rangle]^T, \] where
\(\langle \cdot , \cdot \rangle\) denotes the standard \(\ell_2\) inner
product \citep{shuman_emerging_2013}. The inverse transform is then
given by \[
x = U \hat{x}. \]

It follows that any signal \(\hat{x}\) can be filtered in the spectral
domain by \[ \hat{y} = g_\theta(\Lambda) \hat{x}, \] where the operator
\(g_\theta(\Lambda) \in \R^{M \times M}\), a matrix function, yields a
diagonal matrix of \(M\) Fourier coefficients parametrized by
\(\theta\). The filtering operation can equivalently take place in the
vertex domain as \[ y = U g_\theta(\Lambda) U^T x = g_\theta(\L) x \]
where the operator
\(g_\theta(\L) = g_\theta(U \Lambda U^T) = U g_\theta(\Lambda) U^T \in \R^{M \times M}\)
is akin to a convolution with the filter \(Ug_\theta(\Lambda)\) in the
vertex domain. Note that the filtering operation is defined as a
multiplication in the spectral domain because a convolution cannot be
defined in the vertex domain \citep{shuman_emerging_2013}.

\subsection{Graph Filter Learning}\label{graph-filter-learning}

Given a graph \(G\), a set
\(X = [x_0, \ldots, x_{N-1}] \in \R^{M \times N}\) of \(N\) source graph
signals of dimensionality \(M\) and their associated set
\(Y = [y_0, \ldots, y_{N-1}] \in \R^{M \times N}\) of target signals, we
want to learn the parameters \(\theta \in \R^M\) of a graph filter
\(g_\theta(\Lambda) := \diag(\theta)\) such as to minimize the convex
mean square error \begin{equation} L =
\frac{1}{N} \sum_{i=0}^{N-1} \norm{ g_\theta(\L) x_i - y_i }_2^2 =
\frac{1}{N} \norm{ U \diag(\theta) U^T X - Y }_F^2, \label{eq:loss}\end{equation}
where \(\norm{\cdot}_2^2\) denotes the squared \(\ell_2\) norm and
\(\norm{\cdot}_F^2\) the squared Frobenius norm.

Note that, while being the most flexible definition of
\(g_\theta(\Lambda)\), the independence of the Fourier coefficients
\(\theta\) from the eigenvalues \(\Lambda\) does omit all information
about frequencies. As we know that the lower frequencies are more
important for clustering {[}ref NCut{]}, we could have designed a
parametric filter with less parameters.

Rewriting eq.~\ref{eq:loss} in the spectral domain while decomposing it
w.r.t. the scalar coefficients \(\theta_i\) gives \[ L =
\frac{1}{N} \norm{ \diag(\theta) U^T X - U^T Y }_F^2 =
\frac{1}{N} \sum_{i=0}^{M-1} \norm{\theta_i\Xh_{i,\cdot}-\Yh_{i,\cdot}}_2^2, \]
where \(\Xh = U^TX \in \R^{M \times N}\) and
\(\Yh = U^TY \in \R^{M \times N}\) are the spectral representations of
the signals \(X\) and \(Y\). The gradient for each coefficient is then
given by \[ \nabla_{\theta_i} L =
\frac{2}{N} ( \theta_i \Xh_{i,\cdot} - \Yh_{i,\cdot} ) \Xh^T_{\cdot,i} \]
and can be rewritten in a vectorized form as
\begin{equation} \nabla_{\theta} L =
\frac{2}{N} \diag \left( (\diag(\theta) \Xh - \Yh) \Xh^T \right) =
\frac{2}{N} \left( ( \theta 1_N^T \odot \Xh - \Yh ) \odot \Xh \right) 1_N,
\label{eq:gradient}\end{equation} where \(1_N\) denotes a unit vector of
length \(N\) and \(\odot\) the element-wise Hadamard product. The second
form avoids the computation of the useless off-diagonal elements.

A closed-form solution is given by the optimality condition
\(\nabla_{\theta}L=0\) such that
\begin{equation} \theta^* = \argmin_\theta L =
(\Xh \odot \Yh) 1_N \oslash (\Xh \odot \Xh) 1_N, \label{eq:direct}\end{equation}
where \(\oslash\) denotes an element-wise division. Note that this
method is impractical for large \(M\) and \(N\) (sufficiently large for
\(X\) and \(Y\) to not fit in memory). A \gls{SGD} based on
eq.~\ref{eq:gradient} is then necessary.

There are however two major computational drawbacks to this optimization
process: 1) the Fourier transform \(\Xh=U^TX\) of a set of signals costs
\(O(M^2N)\) operations and 2) the eigenvalue decomposition
\(\L = U \Lambda U^T\) costs \(O(M^3)\). The total cost of filtering is
thus \(O(M^2 \max(M,N))\), a problem already stated in
\citep{henaff_deep_2015}. We propose to overcome these computational
issues using efficient numerical approximations such as the Chebyshev
polynomials or the Lanczos algorithm.

\subsubsection{Chebyshev
Parametrization}\label{chebyshev-parametrization}

The main idea, to avoid the Fourier basis, is twofold: 1) parametrize
the filter, in the spectral domain, as a truncated Chebyshev expansion
and 2) recursively compute the Chebyshev polynomials from the Laplacian.
This approximate method for spectral graph filtering was first proposed
in \citep{hammond_wavelets_2011} for a fast wavelet transform on graphs.

Recall that the Chebyshev polynomial \(T_k(x)\) of order \(k\) may be
generated by the stable recurrence relation
\(T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)\) with \(T_0 = 1\) and
\(T_1 = x\). These polynomials form an orthogonal basis for
\(L^2([-1,1], dy / \sqrt{1-y^2})\), the Hilbert space of square
integrable functions with respect to the measure \(dy/\sqrt{1-y^2}\).

The graph filter \(g_\theta(\Lambda)\) can thus be constructed from the
truncated expansion
\[ g_\theta(\Lambda) := \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\Lambda}) \]
of polynomial order \(K-1\), where the parameter \(\theta \in \R^K\),
\(K \ll M\), is a vector of Chebyshev coefficients and
\(T_k(\tilde{\Lambda}) \in \R^{M \times M}\) is the Chebyshev polynomial
of order \(k\) evaluated at
\(\tilde{\Lambda} := 2 \Lambda / \lambda_{max} - I_M \in \R^{M \times M}\),
a diagonal matrix of scaled eigenvalues (so that they lie in
\([-1,1]\)). While reducing the number of parameters from \(M\) to
\(K\), this parametrization enforces smoothness in the spectral domain,
which translates to localization in the vertex domain. It can indeed be
shown that \((\L^k)_{i,j}=0\) if the shortest-path between vertices
\(v_i\) and \(v_j\) is longer than \(k\) edges
\citep{hammond_wavelets_2011}, which limits the influence of a
\(K^\text{th}\) order filter to \(K\) hopes. This is often a desired
property, e.g.~to learn local features for classification.

To avoid the Fourier basis \(U\), we express the polynomials as
functions of the Laplacian \(\L\) instead of its eigenvalues \(\Lambda\)
using eq.~\ref{eq:lap_diag}, such that the filtering operator in the
vertex domain is given by
\begin{equation} g_\theta(\L) = U g_\theta(\Lambda) U^T =
\sum_{k=0}^{K-1} U \theta_k T_k(\tilde{\Lambda}) U^T =
\sum_{k=0}^{K-1} \theta_k T_k(\tilde{\L}), \label{eq:chebyshev}\end{equation}
where \(T_k(\tilde{\L}) \in \R^{M \times M}\) is the Chebyshev
polynomial of order \(k\) evaluated at the scaled Laplacian
\(\tilde{\L} := 2 \L / \lambda_{max} - I_M\). Note that the spectrum of
the normalized Laplacian is bounded by \(2\)
\citep{chung_spectral_1997}, such that the scaling can simply be
\(\tilde{\L} = \L - I_M\), tolerating some imprecision in the
approximation due to the loss of a fraction \(2-\lambda_{max}\) of the
domain.

Inserting eq.~\ref{eq:chebyshev} into eq.~\ref{eq:loss} we obtain
\begin{equation} L =
\frac{1}{N} \norm{ \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\L}) X - Y }_F^2 =
\frac{1}{N} \norm{ \bar{X} \theta - \bar{y} }_2^2, \label{eq:loss_c}\end{equation}
where \(\bar{y} \in \R^{MN}\) is the vectorized matrix \(Y\) and the
\(k^\text{th}\) column of \(\bar{X} \in \R^{MN \times K}\) is the
vectorized matrix
\(\tilde{X}_k := T_k(\tilde{\L}) X \in \R^{M \times N}\). The gradient
is then given by \begin{equation} \nabla_\theta L =
\frac{2}{N} \bar{X}^T (\bar{X} \theta - \bar{y}). \label{eq:gradient_c}\end{equation}

While the system \(\bar{X} \theta - \bar{y}\) is largely over-determined
as \(K \ll MN\), a closed-form solution of eq.~\ref{eq:loss_c} is given
by the optimality condition \(\nabla_\theta L = 0\) so that
\begin{equation} \theta^* = \argmin_\theta L = \bar{X}^+ \bar{y}, \label{eq:direct_c}\end{equation}
where \(\bar{X}^+ = (\bar{X}^T\bar{X})^{-1} \bar{X}^T\) is the
pseudo-inverse of \(\bar{X}\). The computation of the inverse is fast
and stable for small \(K\). Alternative approximate solution methods for
larger order \(K\) include the computation of the pseudo-inverse with
\gls{SVD}, the use of a least-square solver or a gradient descent scheme
based on eq.~\ref{eq:gradient_c}.

Using the recurrence
\[ \tilde{X}_k = 2\tilde{\L} \tilde{X}_{k-1} - \tilde{X}_{k-2} \] with
\(\tilde{X}_0 = X\) and \(\tilde{X}_1 = \tilde{\L} X\), the computation
of \(\bar{X}\) given \(X\) costs \(O(K|E|N) = O(KMN)\) operations if the
number of edges \(|E|\) is proportional to the number of nodes \(M\),
e.g.~for kNN graphs. As the cost of the product \(\bar{X}\theta\) is
similar, the entire filtering operation has a computational cost of
\(O(KMN)\) whereas the straightforward implementation using the Fourier
basis has a cost of \(O(M^2\max(M,N)\). When applying multiple filters
to the same set of signals, as is the case with \gls{SGD}, one may save
computations by storing and querying \(\bar{X}\) instead of \(X\), at
the expense of \(K\) times the memory.

\subsubsection{Lanczos Parametrization}\label{lanczos-parametrization}

Another applicable parametrization is based on the Lanczos algorithm, an
adaptation of the power iteration to find the largest or smallest
eigenvalues and corresponding eigenvectors of a linear system. It was
first introduced for fast graph filtering in
\citep{susnjara_accelerated_2015}.

The algorithm, described in
\citep{gallopoulos_efficient_1992, susnjara_accelerated_2015},
constructs an orthonormal basis
\(V = [v_0, \ldots, v_{K-1}] \in \R^{M \times K}\) of the Krylov
subspace \(\mathcal{K}_K(\L,x) = \spn\{ x, \L x, \ldots, \L^{K-1} x \}\)
and a tri-diagonal matrix \(H = V^T \L V \in \R^{K \times K}\) with a
computational cost of \(O(K |E|)\). Note that for large
\(K \gtrapprox 30\), the original iterative algorithm may loose the
basis orthogonality such that a necessary orthogonalization step will
increase the complexity \citep{susnjara_accelerated_2015}. Filtering the
signal \(x\) with \(g_\theta(\L)\) can then be approximated by an order
\(K-1\) polynomial as \begin{equation} y = g_\theta(\L) x \approx
V g_\theta(H) V^T x = V Q g_\theta(\Sigma) Q^T V^T x \label{eq:lanczos}\end{equation}
where \(Q \Sigma Q^T = H\) is the eigendecomposition of \(H\). There
exist fast methods for the eigendecomposition of symmetric tri-diagonal
matrices {[}{]}. As for the Chebyshev approximation, this construction
enforces smoothness in the spectral domain.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{color}

\title{Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering}

@@

\begin{abstract}
Convolutional neural networks have greatly improved state-of-the-art performances in computer vision and speech analysis tasks, due to its high ability to extract multiple levels of representations of data. In this work, we are interested in generalizing convolutional neural networks from low-dimensional regular grids, where image, video and speech are represented, to high-dimensional irregular domains, such as social networks, biological graphs like gene regulatory and brain connectivity networks, telecommunication networks, or words' embedding. We present a formulation of convolutional neural networks on graphs in the context of the emerging field of signal processing on graphs, which provides the necessary mathematical background and efficient numerical schemes to design fast localized convolutional filters on graphs. Numerical experiments on MNIST and 20NEWS demonstrate the ability of the system to learn local stationary features on graphs, as long as the graph is well-constructed.
\end{abstract}

%Previous works have been developed along this today essential line of research, but they lack of mathematical foundations and efficient learning of filters.

\begin{figure}[h!]
\centering
\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.eps}
%\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at layer $l$, $G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
\label{fig:}
\end{figure}

\section{Introduction}
\vspace{-0.2cm}
Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST} offer an efficient architecture to extract highly meaningful statistical patterns in large-scale and high-dimensional datasets. The key success of CNNs is its advanced ability to learn local stationary structures and compose them to form multi-scale hierarchical patterns. Precisely, CNNs extract the local stationarity property of the input data or signals by revealing local features that are shared across the data domain. These similar features are identified with localized convolutional filters or kernels, which are learned from the data. Convolutional filters are shift- or translation-invariant filters, meaning they are able to recognize identical features independently of their spatial locations. Localized kernels or compactly supported filters refer to filters that extract local features independently of the input data size, with a support size that can be much smaller than the input size. \\
The compositional local stationarity property of CNNs can be efficiently implemented on regular grids, like image and sound domains, as convolutional and downsampling operators are mathematically well-defined on such discretized Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on low-dimensional regular lattices. But not all data lie on regular grids. User data on social networks, gene data on biological regulatory networks, log data on telecommunication networks, or text documents on words' embedding are important examples of data lying on irregular or non-Euclidean domains. Unfortunately, CNNs cannot be directly applied to such complex domains.\\
A generalization of CNNs to irregular grids is not direct as the basic convolutional and downsampling operators, necessary to extract compositional local stationarity features, are not mathematically clear on these domains. This makes this extension mathematically challenging, both in theory and in implementation. The goal of this work is to leverage the recent works of \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs,art:HammondVandergheynstGribonval11GraphWav,art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG,art:ShumanRicaudVandergheynst16Local} to introduce an efficient implementation of CNNs on irregular domains represented here as graphs. Graphs are universal representations of heterogeneous pairwise relationships between possibly high-dimensional data. Graphs can encode complex geometric structures, and can be studied with strong mathematical tools such as graph spectral theory, a blend between graph theory and harmonic analysis.\\
The major bottleneck of generalizing CNNs to graphs is the design of fast and localized filters on graphs, that are the core elements to extract local stationarity structures. Our major goal is to overcome this barrier. We will make use of recent tools developed in the context of signal processing on graphs (SPG) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to precisely control the spatial support of graph filters using the so-called polynomial kernel localization \cite{art:ShumanRicaudVandergheynst16Local}, and to perform an efficient backward propagation technique using the Chebyshev technique \cite{art:HenaffBrunaLeCun15DLgraphs}. Eventually, a generic and efficient implementation of CNNs on graphs may pave the way to the development of a potentially large class of deep learning techniques for graph-based data.

\section{Related Works}
\vspace{-0.2cm}
{\bf CNNs on non-Euclidean domains.} Extending the success of CNNs to non-Euclidean domains is essential to boost the power analysis of data lying on complex networks like biological, social or telecommunication networks. This is a recent line of works, with a preliminary approach coming from the local reception fields in \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF}, and applied successfully to object recognition in computer vision. The main idea is to unsupervisedly learn groups of similar features. This is highly useful to automatically choose a limited number of connections between two successive layers. This approach is however not able to extract similar groups over the data domain, as it is not clear how to define convolutional filters on these groups.  \\
In \cite{art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, the authors proposed a generalization of CNNs on graphs using graph counterparts of grid convolution and downsampling. They were able to learn shift-invariant or convolutional filters on graphs using the convolution theorem \cite{book:Mallat99wavelets} that defines convolutional operators as linear operators that diagonalize in the Fourier basis (represented by the eigenvectors of the Laplacian operator). They also designed a strategy to learn the graph structure from the data, and apply their model on image recognition, text categorization and bioinformatics. This promising approach is however limited in implementation as forward and backward evaluations have complexity $O(n^2)$ due to the matrix multiplication operations that operate the graph Fourier transforms. Besides, there is no strong control over the support size of the convolutional kernels, which is essential to design localized filters to extract local stationary features, independent of the input size. \\
The authors in \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,art:MasciBoscainiBronsteinVandergheynst15ShapeNet} introduced ShapeNet, a generalization of CNNs to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. They used geodesic polar coordinates to define mesh patches and formulated a deep learning architecture. They obtained state-of-the-art results for shape recognition tasks.\\
Finally, the papers \cite{pro:ChenChengMallat14deepHaar,pro:RustamovGuibas14deepHaar} investigated the construction Haar wavelet transforms on graphs using a deep hierarchical architecture. They applied these deep wavelets on graphs to object recognition on sphere, and to sparse reconstruction of faces. \\
{\bf Signal Processing on Graphs.} SPG is an emerging field that aims at bridging the gap between signal processing techniques like wavelet analysis and graph theory such as spectral graph analysis \cite{art:BelkinNiyogi05LaplaBeltrami,art:VonLuxburg07Tutorial}. The main goal is to generalize fundamental analysis operations for signals from regular grids to irregular structures embodying by graphs. We refer the reader to \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} for an introduction of the field. Standard operations on grids such as convolution, translation, filtering, dilatation, modulation, downsampling do not extend directly to graphs and thus require new mathematical definitions while keeping the original intuitive concepts. In this context, the authors of \cite{art:HammondVandergheynstGribonval11GraphWav,art:CoifmanLafon06DifMap,pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform mutli-scale pyramid transforms on graphs. The works of \cite{pro:TsitsveroBarbarossa15Uncert,pro:PasdeloupAlamiGriponRabbat15Uncert,art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles on graphs, and showed that intuitive concepts may be lost, but can also produce enhanced localization principles for signals on graphs. In \cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to carry out lasso-based signal regularization on graphs, and studied the intertwined relationships between smoothness and sparsity on graphs. In \cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors investigated compressed sensing recovery conditions for graph spectral signal analysis. \\
Beyond this work, our long-term goal is to make use of SPG tools as much as possible to formulate mathematically well-designed techniques of CNNs on graphs. This paper introduces the mathematical and computational foundations of such models, while future developments may consider enhanced graph filters like graph wavelet operators, spectral graph coarsening techniques, and improved localization properties.

\section{Contributions}
\vspace{-0.2cm}
{\textbf {\textit {(1)}}} Learning deep fully connected networks is almost computationally untractable for large data domain. Fortunately, for data satisfying stationarity and hierarchical compositionality structure, CNNs have proven to be remarkably powerful to identify highly informative features using low learning complexity. For example, learning a single layer with $n$ input coordinates and $m<n$ outputs require $O(n^2)$ complexity to learn fully connected networks. It also requires the complexity $O(n)$ to learn any arbitrary filter, and $O(S)$ for localized filters, where $S$ is the size of the filter support, independent of the input size $n$. One of our main contributions is to stay with the same low complexity for learning graph convolutional filters. We will show it is actually possible to keep exactly the same complexity as standard CNNs, i.e. $O(S)$ for localized graph filters.\\
{\textbf {\textit {(2)}}} The backpropagation learning algorithm requires to run convolutional operations, which can be efficiently implemented by FFT on grids with complexity $O(n\log n)$. Convolution on graphs can also be implemented by matrix multiplications representing the forward and backward graph Fourier transforms, but it is expensive as the complexity is $O(n^2)$. As noticed in \cite{art:HenaffBrunaLeCun15DLgraphs}, this complexity is a strong bottleneck to deal with large-scale problems. We will greatly improve this complexity by reducing to $O(E.S)$, where $E$ is the number of edges in the given graph, using the Chebyshev technique from SPG \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. Besides, as most real-world graphs are highly sparse, we have $E\ll n^2$, and therefore a low complexity for the matrix multiplication operations. Eventually, the Chebyshev computations can be naturally distributed on GPUs, as we did in our experiments.\\
{\textbf {\textit {(3)}}} We introduce an efficient pooling strategy for CNNs on graphs, that also makes use of distributed computing architecture like GPUs. The pooling approach consists on two stages. First, a graph coarsening procedure based on METIS \cite{art:DhillonGuanKulis07Graclus,art:KarypisKumar98Metis} is successively applied to the original graph to produce coarser versions of the graph. Second, at each coarsening level, we sequentially order all nodes such that two successive indexed nodes are the parent nodes of the child node in the next coarser level. This node indexation structure in memory satisfies parallel architectures. \\
{\textbf {\textit {(4)}}} We regard this proposed work as the funding paper that bridges the gap between two powerful fields, namely deep learning and signal processing on graphs. In the future, we expect to further benefit from tools in SPG to enhance the formulation of CNNs on graphs. \\
{\textbf {\textit {(5)}}} We revisit the Euclidean case in the numerical experiments with the standard MNIST image classification problem. The model is able to learn localized convolutional filters and achieve 98.65\% of success rate. More interestingly, we apply our model to the text categorization problem on 20NEWS. We also demonstrate that our techniques is able to extract local stationary features, as it improves by 2.5\% the classification accuracy when no convolutional are present. We observe that the quality of the graph construction is strongly linked to the quality of the convolutional layers. Finally, we demonstrate the computational performance of our model over \cite{art:HenaffBrunaLeCun15DLgraphs}.

\section{Learning Localized Graph Filters}
\vspace{-0.2cm}
In this section, we explore how to design efficient localized convolutional filters on graphs.

\subsection{Signal Processing on Graphs and Localized Filters}
\vspace{-0.2cm}
{\bf Graph Fourier Transform.} We are interested in processing signals defined on undirected and connected graphs $G=(V,E,W)$, where $V$ is the set of vertices with $|V|=n$, $E$ is the set of edges and $W$ is the weighted adjacency matrix encoding the connection weights between two vertices. A signal $x:V\rightarrow\mathbb{R}$ defined on the nodes of the graph may be regarded as a vector $x\in\mathbb{R}^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node. An essential operator in spectral graph analysis is the graph Laplacian \cite{book:Chung97Spectral}, which unnormalized definition is $L=D-W\in\mathbb{R}^{n\times n}$ with $D\in\mathbb{R}^{n\times n}$ is the diagonal degree matrix where $D_{ii}=\sum_jW_{ij}$, and its normalized definition is $L=I_n-D^{-1/2}WD^{-1/2}$ where $I_n$ is the identity matrix. As $L$ is a real symmetric positive semidefinite matrix, it has a complete set of orthonormal eigenvectors $\{u_l\}_{l=0}^{n-1}\in\mathbb{R}^n$, known as the graph Fourier modes, and their associated real nonnegative eigenvalues $\{\lambda_l\}_{l=0}^{n-1}$, which are the frequencies of the graph system. Equivalently, linear operator $L$ diagonalizes the Fourier basis matrix $U=[u_0,...,u_{n-1}]\in\mathbb{R}^{n\times n}$ such that $L=U\Lambda U^T$ where $\Lambda=\diag(\lambda_l)$. We can now define the graph Fourier transform (GFT) and its inverse \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. The GFT of a spatial signal $x\in\mathbb{R}^n$ is $\hat{x}=U^Tx\in\mathbb{R}^n$, and the inverse GFT is $x=U\hat{x}$. Using GFT we can now generate fundamental operations of signal processing on graphs such as convolution and filtering. \\
{\bf Convolutional Filters.} The graph convolutional operator is simply defined in the Fourier space such as $f\ast_G g=U((U^Tf) \odot (U^Tg))$, where $\odot$ the element-wise Hadamard product. Then the convolutional filtering operation on graphs is defined as $y=g_\theta(L)x\in\mathbb{R}^n$, where $g_\theta(L)=g_\theta(U\Lambda U^T)=Ug_\theta(\Lambda) U^T\in\mathbb{R}^{n\times n}$ where $g_\theta(\Lambda)=\diag(\theta)$ and $\theta\in\mathbb{R}^n$ is the filter expressed in the Fourier space. Remember that convolutional filters, i.e. translation-invariant filters on graphs, can only be defined in the Fourier domain as convolution operators are not defined in the vertex space. \\
{\bf Localized Convolutional Filters.} Learning convolutional filters on graphs boils down to learn spectral multipliers $\theta$ of complexity $O(n)$, the size of the input data. Our goal is to reduce this complexity to $O(S)$, the support size of the filter in the vertex space, exactly as in standard CNNs. For this, we will make use of a property called polynomial localization \cite{art:ShumanRicaudVandergheynst16Local}. Let us represent $\theta=\{\theta_{\{{\lambda_l}\}_{l=0}^n}\}\in\mathbb{R}^n$ as a $K$-th differentiable polynomial function: $\theta_{\lambda_l}^K(a)=\sum_{k=0}^{k-1}a_k \lambda_l^k$, where $a=\{a_k\}_{k=0}^{K-1}\in\mathbb{R}^K$. The polynomial localization theorem \cite{art:ShumanRicaudVandergheynst16Local} shows that $(T_ig_{\theta^K})_j=(\delta_i \ast_G (Ug_{\theta^K}))_j=0$ for $d_G(i,j)>K$ where $d_G$ is the shortest distance on graph, meaning that the support of the associated spatial filter is exactly $K$-localized, all values are zero after $K$ hopes. Consequently, spectral filters represented by $K$-order polynomials are spatially localized with support $S=K$, implying the low learning complexity $O(K)$ for each filter. We notice that the authors \cite{art:HenaffBrunaLeCun15DLgraphs} also used a $K$-order polynomials represented by splines, but they did not formalize the quantitive relationship between filter support size and polynomial order. Finally, it is also possible to estimate the spatial decaying profile of the filter \cite{art:ShumanRicaudVandergheynst16Local}: for any K-localized filter centered at vertex $i$, we have the profile $|(T_ig_{\theta^K})(j)|\leq\big[\frac{2\sqrt{n}}{K!}(\frac{\lambda_{max}}{4})^K \sup_{\lambda}|(Ug_{\theta^K})^K(\lambda)|\big]$, which may be useful for graph filter analysis.
